### 2017-04-27 article/article.tex
Started Repo

% !TEX root = ./article.tex

\documentclass{article}

\usepackage{mystyle}
\usepackage{myvars}

%-----------------------------

\begin{document}

	\maketitle % Insert title

	\thispagestyle{fancy} % All pages have headers and footers

%-----------------------------
%	ABSTRACT
%-----------------------------

	\begin{abstract}
		\noindent [TODO ]
	\end{abstract}

%-----------------------------
%	TEXT
%-----------------------------

	\section{[TODO ]}
	\label{sec:e1}

		\paragraph{}
		[TODO ]

%-----------------------------
%	Bibliographic references
%-----------------------------
	\nocite{garciparedes:machine-learning-bayesian-2}
	\nocite{subject:taa}
	\nocite{tool:weka}
  \bibliographystyle{alpha}
  \bibliography{bib/misc}

\end{document}